---
title: "Weather and Corn Yield Regressions"
author: "Alex Siggers"
date: "03/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(R.matlab)
library(rnassqs)
library(ggplot2)
```


```{r tmax data, echo=FALSE, include=FALSE, message=FALSE, warning=FALSE}
# daily max temperature
# dimensions: counties x days x years
prism <- readMat("data/prismiowa.mat")

# look at county #1
t_1981_c1 <- prism$tmaxdaily.iowa[1,,1]
t_1981_c1[366]
plot(1:366, t_1981_c1, type = "l")
ggplot() +
  geom_line(mapping = aes(x=1:366, y = t_1981_c1)) +
  theme_bw() +
  xlab("day of year") +
  ylab("daily maximum temperature (°C)") +
  ggtitle("Daily Maximum Temperature, Iowa County #1")

```


```{r tidying up, echo=FALSE, include=FALSE, message=FALSE}
# assign dimension names to tmax matrix
dimnames(prism$tmaxdaily.iowa) <- list(prism$COUNTYFP, 1:366, prism$years)
# converted 3d matrix into a data frame
tmaxdf <- as.data.frame.table(prism$tmaxdaily.iowa)
# relabel the columns
colnames(tmaxdf) <- c("countyfp","doy","year","tmax")
tmaxdf <- tibble(tmaxdf)
```

```{r temp trends, echo=FALSE, include=FALSE, message=FALSE}
tmaxdf$doy <- as.numeric(tmaxdf$doy)
tmaxdf$year <- as.numeric(as.character(tmaxdf$year))
winnesummer <- tmaxdf %>%
  filter(countyfp==191 & doy >= 152 & doy <= 243) %>%
  group_by(year) %>%
  summarize(meantmax = mean(tmax))
ggplot(winnesummer, mapping = aes(x = year, y = meantmax)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)") +
  geom_smooth(method = lm)
lm_summertmax <- lm(meantmax ~ year, winnesummer)
summary(lm_summertmax)
```


```{r winter temps, echo=FALSE, include=FALSE, message=FALSE}
winnewinter <- tmaxdf %>%
  filter(countyfp==191 & doy <= 59 | doy >= 335 & !is.na(tmax)) %>%
  group_by(year) %>%
  summarize(meantmax = mean(tmax))
ggplot(winnewinter, mapping = aes(x = year, y = meantmax)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)") +
  geom_smooth(method = lm)
lm_wintertmax <- lm(meantmax ~ year, winnewinter)
summary(lm_wintertmax)
```



```{r quadratic temp trend, echo=FALSE, include=FALSE, message=FALSE}
winnewinter$yearsq <- winnewinter$year^2
lm_wintertmaxquad <- lm(meantmax ~ year + yearsq, winnewinter)
summary(lm_wintertmaxquad)
winnewinter$fitted <- lm_wintertmaxquad$fitted.values
ggplot(winnewinter) +
  geom_point(mapping = aes(x = year, y = meantmax)) +
  geom_line(mapping = aes(x = year, y = fitted)) +
  theme_bw() +
  labs(x = "year", y = "tmax")
```


```{r yield download, echo=FALSE, include=FALSE, message=FALSE}
# set our API key with NASS
nassqs_auth(key = "9DB94D80-F55F-379D-AF64-B3EC20632EB0")

# parameters to query on 
params <- list(commodity_desc = "CORN", util_practice_desc = "GRAIN", prodn_practice_desc = "ALL PRODUCTION PRACTICES", year_GE = 1981, state_alpha = "IA")

# download
cornyieldsall <- nassqs_yields(params)

cornyieldsall$county_ansi <- as.numeric(cornyieldsall$county_ansi)
cornyieldsall$yield <- as.numeric(cornyieldsall$Value)

# clean and filter this dataset
cornyields <- select(cornyieldsall, county_ansi, county_name, yield, year) %>%
  filter(!is.na(county_ansi) & !is.na(yield))
cornyields <- tibble(cornyields)

view(cornyields)
```

## Assignment

### Question 1a: Extract Winneshiek County corn yields, fit a linear time trend, make a plot. Is there a significant time trend?

```{r, echo=TRUE, message=FALSE, warning=FALSE}
#1a
WinneshiekCorn <- cornyields %>%
  filter(county_name == "WINNESHIEK")

ggplot(WinneshiekCorn, mapping = aes(x = year, y = yield)) +
  geom_point() +
  theme_bw() +
  labs(x = "Year", y = "Corn Yield") +
  geom_smooth(method = lm)

WinCornLm <- lm(yield ~ year, WinneshiekCorn)
summary(WinCornLm)
```

-##There is a very significant linear increase in corn yield with year. The linear model provides evidence, and the scatterplot visualizes the trend. ##


### Question 1b: Fit a quadratic time trend (i.e., year + year^2) and make a plot. Is there evidence for slowing yield growth? 

```{r, echo=TRUE, message=FALSE, warning=FALSE}
#1b
WinneshiekCorn$yearsq <- WinneshiekCorn$year^2

lm_wincornquad <- lm(yield ~ year + yearsq, WinneshiekCorn)
summary(lm_wincornquad)

WinneshiekCorn$fitted <- lm_wincornquad$fitted.values

ggplot(WinneshiekCorn) +
  geom_point(mapping = aes(x = year, y = yield)) +
  geom_line(mapping = aes(x = year, y = fitted)) +
  theme_bw() +
  labs(x = "Year", y = "Corn Yield")
```

-##The evidence seems to suggest that the corn yield is only growing, seemingly exponentially under the quadratic fit. ##


### Question 2 -- Time Series: Let's analyze the relationship between temperature and yields for the Winneshiek County time series. Use data on yield and summer avg Tmax. Is adding year or Tmax^2 to your model helpful? Make a plot and interpret the results.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
WinTime = merge(WinneshiekCorn, winnesummer, by="year")

YieldTemplm=lm(yield~meantmax, WinTime)
summary(YieldTemplm)

ggplot(WinTime, mapping = aes(x = meantmax, y = yield)) +
  geom_point() +
  theme_bw() +
  labs(x = "Average Summer Temp", y = "Corn Yield") +
  geom_smooth(method = lm)
```

-##Visually, there appears to be a negative correlation between average maximum temperature and yield, but the model does not provide strong evidence for such a relationship. Adding year to the model provides significant evidence for a relationship in that regard, but we are note exactly looking for that relationship here. ##


### Question 3 -- Cross-Section: Analyze the relationship between temperature and yield across all counties in 2018. Is there a relationship? Interpret the results.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
#Q3
rename(tmaxdf, "county_ansi" = "countyfp")
names(tmaxdf)[names(tmaxdf)=="countyfp"] <- "county_ansi"

TempData2018 <- tmaxdf %>%
  filter(doy >= 152 & doy <= 243) %>%
  filter(year == "2018") %>%
  group_by(county_ansi)

Data2018 <- cornyields %>%
  filter(year == "2018")
  
TD2018 <- merge(TempData2018, Data2018, by="county_ansi")
  
  

lm2018 <- lm(yield~tmax, TD2018)
summary(lm2018)

ggplot(TD2018, mapping = aes(x = tmax, y = yield)) +
  geom_point() +
  theme_bw() +
  labs(x = "Max Summer Temp", y = "Corn Yield") +
  geom_smooth(method = lm)

```

-##These results appear to mimic the trend of the summer Winnishiek regression, which is minimal. There is a slight decreasing trend, but overall very little evidence of a trend at all. ##

### Question 4 -- Panel: One way to leverage multiple time series is to group all data into what is called a "panel" regression. Convert the county ID code ("countyfp" or "county_ansi") into factor using as.factor, then include this variable in a regression using all counties' yield and summer temperature data. How does the significance of your temperature coefficients (Tmax, Tmax^2) change? Make a plot comparing actual and fitted yields and interpret the results of your model.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(gridExtra)
#Q4
#Converting to factors
cornyields$county_ansi=as.factor(cornyields$county_ansi)
cornyields$county_name=as.factor(cornyields$county_name)
tmaxdf$county_ansi=as.factor(tmaxdf$county_ansi)

#Creating tmax and tmax^2 df
MaxTemps <- tmaxdf %>%
  filter(doy >= 152 & doy <= 243) %>%
  group_by(year, county_ansi) %>%
  summarize(meantmax = mean(tmax))

MaxTemps$county_ansi=as.factor(MaxTemps$county_ansi)

#Creating master df
MaxTempYield = merge(MaxTemps, cornyields) %>%
  mutate(yearsq=year^2,
         yieldsq=yield^2,
         tmaxsq=meantmax^2)

#Creating models for each factor
tempquadr = lm(yield~meantmax + tmaxsq, MaxTempYield)
summary(tempquadr)
MaxTempYield$tempsqfit <- tempquadr$fitted.values

yearlm = lm(yield~year, MaxTempYield)
summary(yearlm)
MaxTempYield$yearfit <- yearlm$fitted.values

countylm = lm(yield~county_name, MaxTempYield)
summary(countylm)
MaxTempYield$countyfit <- countylm$fitted.values

allquadr = lm(yield ~ year + meantmax + tmaxsq + county_name, MaxTempYield)
summary(allquadr)
MaxTempYield$allfit <- allquadr$fitted.values

summary(MaxTempYield)
```


```{r, echo=TRUE, message=FALSE, warning=FALSE}
#Creating visuals'
grid.arrange(
top = "Individual factors regressed against yield for all counties",
layout_matrix = rbind(c(1, 2, 3),c(4, 4, 4)),
ggplot(MaxTempYield) +
  geom_point(mapping = aes(x = yield, y = tempsqfit), alpha = 0.2) +
  geom_line(mapping = aes(x = yield, y = yield), linetype = "solid", size=0.5) +
  theme_bw() +
  labs(x = "Yield (Actual)", y = "Yield (Fitted)", title= "Quad Temp",)
,
ggplot(MaxTempYield) +
  geom_point(mapping = aes(x = yield, y = yearfit), alpha = 0.2) + 
  geom_line(mapping = aes(x = yield, y = yield), linetype = "solid", size=0.5) +
  theme_bw() +
  labs(x = "Yield (Actual)", y = "Yield (Fitted)", title= "Year")
,
ggplot(MaxTempYield) +
  geom_point(mapping = aes(x = yield, y = countyfit), alpha = 0.2) + 
  geom_line(mapping = aes(x = yield, y = yield), linetype = "solid", size=0.5) +
  theme_bw() +
  labs(x = "Yield (Actual)", y = "Yield (Fitted)", title= "County Model")
,
ggplot(MaxTempYield) +
  geom_point(mapping = aes(x = yield, y = allfit), alpha = 0.2) + 
  geom_line(mapping = aes(x = yield, y = yield), linetype = "solid", size=1.5) +
  theme_bw() +
  labs(x = "Yield (Actual)", y = "Yield (Fitted)", title= "Combined Model VS Actual", subtitle="Quadratic Temp, Year, County")
)

```

-##The fit of years to yield (although not quadratic) provides the best fit to explain the correlation. The temperature and county models have very limited predictive power (though the temperature model is better than the county model). The combined model creates the best prediction visually. ##

### Question 5 -- Soybeans: Download NASS data on soybean yields and explore either a time series relationship for a given county, the cross-sectional relationship for a given year, or a panel across all counties and years.

```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#Reading in & downloading data
soypar <- list(
  commodity_desc = "SOYBEANS", statisticcat_desc= "YIELD", 
  prodn_practice_desc = "ALL PRODUCTION PRACTICES", year__GE = 1981, state_alpha = "IA")

soyyield<- nassqs_yields(soypar)
```


```{r, echo=TRUE, message=FALSE, warning=FALSE}
soyyield$county_ansi <- as.numeric(soyyield$county_ansi)
soyyield$yield <- as.numeric(soyyield$Value)

# clean and filter this dataset
soyyields <- select(soyyield, county_ansi, county_name, yield, year) %>%
  filter(!is.na(county_ansi) & !is.na(yield))
soyyields <- tibble(soyyields)


#Subsetting a single county
CherokeeSoy <- soyyields %>%
  filter(county_name == "CHEROKEE")

ggplot(CherokeeSoy, mapping = aes(x = year, y = yield)) +
  geom_point() +
  theme_bw() +
  labs(x = "Year", y = "Soy Yield") +
  geom_smooth(method = lm)

CherSoyLm <- lm(yield ~ year, CherokeeSoy)
summary(CherSoyLm)

```

-There is a very evident positive correlation between time and soy yield in Cherokee county, which is supported by the linear model as well.

### Bonus: Find a package to make a county map of Iowa displaying some sort of information about yields or weather. Interpret your map.

### Bonus #2: Challenge question - map trends in corn yields by county across Iowa. Interpret your map.
